---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# actverse

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![R-CMD-check](https://github.com/gipso/actverse/workflows/R-CMD-check/badge.svg)](https://github.com/gipso/actverse/actions)
[![Codecov test coverage](https://codecov.io/gh/gipso/actverse/branch/main/graph/badge.svg)](https://app.codecov.io/gh/gipso/actverse?branch=main)
[![License: MIT](https://img.shields.io/badge/license-MIT-green)](https://choosealicense.com/licenses/mit/)
[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-v2.0%20adopted-ff69b4.svg)](https://gipso.github.io/actverse/CODE_OF_CONDUCT.html)
<!-- badges: end -->

## Overview

`actverse` is an R package that provides a complete and consistent toolkit to process, analyze and visualize actigraphy data in R. The aim of `actverse` is to facilitate the work of sleep and chronobiology scientists with actigraphy data while also helping with research reproducibility.

> Please note that this package is currently on the development stage and have not yet been [peer reviewed](https://devguide.ropensci.org/softwarereviewintro.html).

## Prerequisites

You only need to have some familiarity with the [R programming language](https://www.r-project.org/) to use the `actverse` main functions.

In case you don't feel comfortable with R, we strongly recommend checking Hadley Wickham and Garrett Grolemund's free and online book [R for Data Science](https://r4ds.had.co.nz/) and the Coursera course from the John Hopkins University [Data Science: Foundations using R](https://www.coursera.org/specializations/data-science-foundations-r) (free for audit students).

## Installation

You can install the development version from GitHub with:

``` {r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("gipso/actverse")
```

## Citation

If you use `actverse` in your research, please consider citing it. We put a lot of work to build and maintain a free and open-source R package. You can find the `actverse` citation below.

```{r}
citation("actverse")
```

## Contributing

We welcome contributions, including bug reports. Take a moment to review our [Guidelines for Contributing](https://gipso.github.io/actverse/CONTRIBUTING.html).

Please note that `actverse` is released with a [Contributor Code of Conduct](https://gipso.github.io/actverse/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
