---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# actverse

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![R-CMD-check](https://github.com/gipso/actverse/workflows/R-CMD-check/badge.svg)](https://github.com/gipso/actverse/actions)
[![Codecov test coverage](https://codecov.io/gh/gipso/actverse/branch/main/graph/badge.svg)](https://app.codecov.io/gh/gipso/actverse?branch=main)
[![License: MIT](https://img.shields.io/badge/license-MIT-green)](https://choosealicense.com/licenses/mit/)
[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-v2.0%20adopted-ff69b4.svg)](https://gipso.github.io/actverse/CODE_OF_CONDUCT.html)
<!-- badges: end -->

## Overview

`actverse` is an R package that provides a complete and consistent toolkit to process, analyze and visualize actigraphy data in R. The aim of `actverse` is to facilitate the work of sleep and chronobiology scientists with actigraphy data while also helping with research reproducibility.

`actverse` adheres to the [tidyverse principles](https://tidyverse.tidyverse.org/articles/manifesto.html) and integrates with the [tidyverse ecosystem](https://www.tidyverse.org/). That's why we choose to work with the [`tsibble`](https://tsibble.tidyverts.org/index.html) standard for time series, that also follow the same principles.

## Prerequisites

You only need to have some familiarity with the [R programming language](https://www.r-project.org/) to use the `actverse` main functions.

In case you don't feel comfortable with R, we strongly recommend checking Hadley Wickham and Garrett Grolemund's free and online book [R for Data Science](https://r4ds.had.co.nz/) and the Coursera course from the John Hopkins University [Data Science: Foundations using R](https://www.coursera.org/specializations/data-science-foundations-r) (free for audit students).

## Installation

You can install the development version from GitHub with:

``` {r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("gipso/actverse")
```

## Usage

Our process of development is continuous and based on an agile approach. That 
means that you can always count with new features on the development version without the need to wait for the next published version.

Our plan is to launch a first version of `actverse` in the month of August of 2022. Until them, you can install the development version and use the features already available (see the Installation section).

### Time series standard

The R ecosystem has a [vast number](https://cran.r-project.org/view=TimeSeries) of time series standards and we had to use one of them. A standard for time objects is really a must, since time can have many representations and can be rooted in  different numerical systems.

We found that the best standard available for packages that adheres the [tidyverse principles](https://tidyverse.tidyverse.org/articles/manifesto.html) is the [`tsibble`](https://tsibble.tidyverts.org/index.html). That way, all of the `actverse` functions will require that your data is in this format. Please refer to [`tsibble`](https://tsibble.tidyverts.org/index.html) documentation to learn how to format your data.

We also recommend seeing the [`tsbox`](https://www.tsbox.help/) package, an R
package that propose to be an "universal translator" (🖖) for R time series standards.

## Read/Write

### Period functions

### Data interpolation

### API clients

### Other features

`actverse()` also comes equipped with many utility functions. The package also provides free datasets for testing and learning purposes.

All functions are well documented, showing all the guidelines behind the computations. Click [here](http://gipso.github.io/actverse) to see a list of them.

## Citation

If you use `actverse` in your research, please consider citing it. We put a lot of work to build and maintain a free and open-source R package. You can find the `actverse` citation below.

```{r}
citation("actverse")
```

## Contributing

We welcome contributions, including bug reports. Take a moment to review our [Guidelines for Contributing](https://gipso.github.io/actverse/CONTRIBUTING.html).

Please note that `actverse` is released with a [Contributor Code of Conduct](https://gipso.github.io/actverse/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
